### 2025-01-29 scripts/Paper-analysis.Rmd
more analysis

# Order the PTUs by mean values of depletion for better visual display
mean.values = diff.leading.region.average %>% group_by(PTU) %>%
  summarise(mean=mean(diff.density)*1000)
diff.leading.region.average$PTU = ordered(diff.leading.region.average$PTU,
                                          levels=mean.values$PTU[order(mean.values$mean)])
p.leading.region.diff.density = ggplot(diff.leading.region.average, aes(as.factor(PTU), diff.density*1000))+ # depletion per-kb rather than per-bp
  ggbeeswarm::geom_quasirandom(alpha=0.8)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  geom_hline(yintercept = 0)+
  xlab("")+
  theme(axis.text=element_text(colour="black"))+
  ylab("Palindrome depletion in leading region (per kb)")+
  coord_flip()
p.leading.region.diff.density
# Based on this, remove two outliers
p.leading.region.diff.density = p.leading.region.diff.density+
  ylim(c(-10,5))
ggsave(p.leading.region.diff.density, file='../results/6mer-deviation-leading-region-vs-rest-of-plasmid-density.pdf',
       width=5, height=4)
```

>>>>

# Order the PTUs by mean values of depletion for better visual display
mean.values = diff.leading.region.average %>% group_by(PTU) %>%
  summarise(mean=mean(diff.density)*1000)
diff.leading.region.average$PTU = ordered(gsub("PTU-", "", diff.leading.region.average$PTU),
                                          levels=gsub("PTU-", "", mean.values$PTU[order(mean.values$mean)]))
p.leading.region.diff.density = ggplot(diff.leading.region.average, aes(as.factor(PTU), diff.density*1000))+ # depletion per-kb rather than per-bp
  ggbeeswarm::geom_quasirandom(alpha=0.8)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  geom_hline(yintercept = 0)+
  xlab("")+
  theme(axis.text=element_text(colour="black", size=18))+
  ylab("Palindrome depletion in leading region (per kb)")+
  theme(axis.text.x=element_text(angle=45, hjust=1))+
  ylab("")
p.leading.region.diff.density
# Based on this, remove two outliers
p.leading.region.diff.density = p.leading.region.diff.density+
  scale_y_continuous(breaks=seq(-8, 4, 2), limits=c(-8,5))
ggsave(p.leading.region.diff.density, file='../results/6mer-deviation-leading-region-vs-rest-of-plasmid-density.pdf',
       width=5.5, height=6)
```
### Averaging over all plasmids

Similar to Samuel et al.'s display of anti-defence gene density, average over all conjugative plasmids.

```{r}
overall.median = d.reduced %>% filter(x<50000) %>%
  group_by(x) %>%
  summarise(y=median(diff_density),
            n=length(diff_density))
p.overall.average = ggplot(overall.median, aes(x, y*1000))+
  geom_line()+
  scale_y_continuous(breaks=seq(-10, 0, 2), limits = c(-9, 0))+
  theme_bw()+
  theme(panel.grid = element_blank())+
  geom_hline(yintercept = 0, linetype='dashed')+
  ylab("Palindrome depletion (per kb)")+
    scale_x_continuous(labels = function(x) x / 1000) +
  xlab("Position on plasmid (kb)")+
  theme(axis.text=element_text(colour="black"))
p.overall.average
# Save
ggsave(p.overall.average, file='../results/6mer-deviation-overall.pdf', 
       width=3, height=3)
```

```{r}
lagging.df = read.csv('~/Downloads/trieste/leading_lagging/NZ_LT994833.1_lagging_10000_k6.csv', header=T)
leading.df = read.csv('~/Downloads/trieste/leading_lagging/NZ_LT994833.1_leading_10000_k6.csv', header=T)
merged.df = merge(leading.df, lagging.df, by="word")
palindromes = tolower(read.csv('~/Downloads/trieste/k6.txt', header=F)$V1)
ggplot(merged.df[which(merged.df$word %in% palindromes),], aes(score.x, score.y))+
  geom_point()
merged.df.palindromes = merged.df[which(merged.df$word %in% palindromes),]
table(merged.df.palindromes$score.x<merged.df.palindromes$score.y)
kmer_targets = tolower(read.csv('~/Downloads/trieste/PTU-LM_targets.txt', header=F)$V1)
ggplot(merged.df[which(merged.df$word %in% kmer_targets),], aes(score.x, score.y))+
  geom_point()
merged.df.targets = merged.df[which(merged.df$word %in% kmer_targets),]
table(merged.df.targets$score.x<merged.df.targets$score.y)
